---
output: html_document
---

### Practical Machine Learning -- Course Project

##### Background

In this project, our goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 

The training data for this project are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

##### Pre-processing

Load libraries.
```{r}
library(caret)
```

Set seed for reproducible results.
```{r}
set.seed(1982)
```

Load the data.
```{r, cache = TRUE}
train <- read.csv("pml-training.csv", na.strings = c("NA", ""))
test <- read.csv("pml-testing.csv", na.strings = c("NA", ""))
```

Clean the data.
```{r}
# Remove columns that are empty in the test set from both train and test sets.
good <- colSums(is.na(test)) == 0
train <- train[ , good]
test <- test[ , good]
```

Remove unuseful columns.
```{r}
train <- train[ , -grep("user_name|X|timestamp|window", names(train))]
test <- test[ , -grep("user_name|X|timestamp|window", names(test))]
```

Split the train set: 80% training and 20% testing.
```{r}
inTrain <- createDataPartition(y = train$classe, p = 0.80, list = FALSE)
training <- train[inTrain, ]
testing <- train[-inTrain, ]
```

Remove columns whose variables contribute to near zero of the variance.
```{r}
nearZeroVar(training)
# There are none to remove.
```


##### Model Fitting

Fit model using simple cross-validation and random forest method with train().
```{r, cache = TRUE}
mycontrol <- trainControl(method = "cv", number = 5)
fit <- train(classe ~ ., data = training, method = "rf", trControl = mycontrol, 
             verbose = TRUE)
```

Results.
```{r}
pred  <- predict(fit, testing) 
confusionMatrix(pred, testing$classe)
# Out of sample error = 1 - Accuracy = 0.48%
(1 - 0.9952)*100
```


##### Predictions

Run the model fit on the test set.
```{r}
answers <- predict(fit, test)
answers
```